---
title: "AI and Data Science Salaries"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cosmo
    css: custom.css
    logo: "C:/Users/patri/Data Visualization Assignment 4/putLogoColor.png" 
runtime: shiny
---


```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(plotly)
library(DT)
library(dplyr)
library(leaflet)

# Load dataset
salaries <- read.csv("salaries.csv")
salaries <- na.omit(salaries)

# Convert relevant columns to factors
salaries$experience_level <- factor(salaries$experience_level, levels = c("EN", "MI", "SE", "EX"))
salaries$employment_type <- factor(salaries$employment_type, levels = c("PT", "FT", "CT", "FL"))
salaries$remote_ratio <- factor(salaries$remote_ratio)
salaries$company_size <- factor(salaries$company_size, levels = c("S", "M", "L"))

# Define country coordinates for the map (You can expand this list)
country_coords <- data.frame(
  country = c('US', 'GB', 'CA', 'IE', 'TR', 'IL', 'IN', 'FI', 'DE', 'LU', 'ZA',
              'KE', 'RS', 'PT', 'AR', 'GR', 'NZ', 'UA', 'LT', 'AU', 'FR', 'ES',
              'SE', 'MX', 'PL', 'NL', 'DK', 'LV', 'AT', 'BR', 'CH', 'AE', 'SA',
              'EG', 'PH', 'OM', 'BA', 'IT', 'AS', 'EE', 'MT', 'HU', 'LB', 'RO',
              'VN', 'NG', 'GI', 'CO', 'SI', 'MU', 'RU', 'KR', 'CZ', 'QA', 'GH',
              'AD', 'EC', 'NO', 'JP', 'HK', 'CF', 'SG', 'TH', 'HR', 'AM', 'PK',
              'IR', 'BS', 'PR', 'BE', 'ID', 'MY', 'HN', 'DZ', 'IQ', 'CN', 'CL',
              'MD'),
  lat = c(37.0902, 55.3781, 56.1304, 53.1424, 38.9637, 31.0461, 20.5937, 61.9241, 51.1657, 49.8153, -30.5595,
          -1.2921, 44.0165, 39.3999, -38.4161, 39.0742, -40.9006, 48.3794, 55.1694, -25.2744, 46.6034, 40.4637,
          60.1282, 23.6345, 51.9194, 52.1326, 56.2639, 56.8796, 47.5162, -14.2350, 46.8182, 23.4241, 23.8859,
          26.8206, 12.8797, 21.5126, 43.9159, 41.8719, -14.2710, 58.5953, 35.9375, 47.1625, 33.8547, 45.9432,
          14.0583, 9.081999, 36.1408, 4.5709, 46.1512, -20.3484, 61.5240, 35.9078, 49.8175, 25.3548, 7.9465,
          42.5462, -1.8312, 60.4720, 36.2048, 22.3193, 6.6111, 1.3521, 15.8700, 45.1, 40.0691, 30.3753,
          32.4279, 25.0343, 18.2208, 50.8503, -0.7893, 4.2105, 15.1990, 28.0339, 33.2232, 35.8617, -35.6751,
          47.4116),
  lon = c(-95.7129, -3.4360, -106.3468, -7.6921, 35.2433, 34.8516, 78.9629, 25.7482, 10.4515, 6.1296, 22.9375,
          36.8219, 21.0059, -8.2245, -63.6167, 21.8243, 174.8859, 31.1656, 23.8813, 133.7751, 2.2137, -3.7492,
          18.6435, -102.5528, 19.1451, 5.2913, 9.5018, 24.6032, 14.5501, -51.9253, 8.2275, 53.8478, 45.0792,
          30.8025, 121.7740, 55.9233, 17.6791, 12.5674, -170.1322, 25.0136, 14.3754, 19.5033, 35.8623, 24.9668,
          108.2772, 8.6753, -5.3536, -74.2973, 14.9955, 57.5522, 105.3188, 127.7669, 15.4729, 51.1839, -1.0232,
          1.6016, -78.1834, 8.4689, 138.2529, 114.1694, 20.9394, 103.8198, 100.9925, 15.2, 45.0382, 69.3451,
          53.6880, -77.3963, -66.5901, 4.3517, 113.9213, 101.9758, -86.2419, 1.6596, 43.6793, 104.1954, -71.5429,
          28.3699)
)

```

Overview
=======================================================================

Column {data-width=350}
-------------------------------------


```{r}
selectInput("experience", "Experience Level:", 
            choices = levels(salaries$experience_level), 
            selected = "SE")
```

```{r}
selectInput("employment", "Employment Type:", 
            choices = levels(salaries$employment_type), 
            selected = "FT")
```

```{r}
selectInput("remote", "Remote Work Ratio:", 
            choices = levels(salaries$remote_ratio), 
            selected = "100")
```

```{r}
selectInput("company_size", "Company Size:", 
            choices = levels(salaries$company_size), 
            selected = "L")

```


Column {data-width=650}
-------------------------------------

### Salary Distribution

```{r}

renderPlotly({
  filtered_data <- salaries %>% 
    filter(experience_level == input$experience,
           employment_type == input$employment,
           remote_ratio == input$remote,
           company_size == input$company_size)
  
  p <- ggplot(filtered_data, aes(x=salary_in_usd)) +
    geom_histogram(binwidth=5000, fill="#0073C2FF", color="white") +
    theme_minimal() +
    labs(title="Distribution of Salaries in USD", x="Salary (USD)", y="Count")
  
  ggplotly(p)
})

```


Table
=======================================================================
```{r}
renderDataTable({
  datatable(salaries, selection = "single", options = list(pageLength = 10, autoWidth = TRUE))
})

```

Experience Level
=======================================================================
```{r}
renderPlotly({
  p <- ggplot(salaries, aes(x=experience_level, y=salary_in_usd, fill=experience_level)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title="Salary by Experience Level", x="Experience Level", y="Salary (USD)") +
    scale_fill_brewer(palette="Set3")
  
  ggplotly(p)
})
```


Employment Type
=======================================================================
```{r}
renderPlotly({
  p <- ggplot(salaries, aes(x=employment_type, y=salary_in_usd, fill=employment_type)) +
    geom_bar(stat="summary", fun="mean") +
    theme_minimal() +
    labs(title="Average Salary by Employment Type", x="Employment Type", y="Average Salary (USD)") +
    scale_fill_brewer(palette="Set2")
  
  ggplotly(p)
})

```


Salary Distribution
=======================================================================

Column {data-width=350}
-------------------------------------

```{r}
selectInput("job_title", "Job Title:", 
            choices = unique(salaries$job_title), 
            selected = unique(salaries$job_title)[1])

```

Column {data-width=1000}
-------------------------------------
```{r}
renderPlotly({
  filtered_data <- salaries %>% 
    filter(job_title == input$job_title)
  
  avg_salaries <- filtered_data %>%
    group_by(company_location) %>%
    summarise(avg_salary = mean(salary_in_usd, na.rm = TRUE)) %>%
    arrange(desc(avg_salary)) %>%
    slice_head(n = 10)
  
  top_countries_data <- filtered_data %>%
    filter(company_location %in% avg_salaries$company_location)
  
  p <- ggplot(top_countries_data, aes(x=company_location, y=salary_in_usd, color=company_location)) +
    geom_point() +
    geom_line(aes(group=company_location)) +
    theme_minimal() +
    labs(title="Top 10 Countries with Highest Salaries", x="Country", y="Salary (USD)") +
    theme(legend.title = element_blank())
  
  ggplotly(p)
})

```

Column {data-width=350}
-------------------------------------
Average Salary

```{r}
renderGauge({
  avg_salary <- salaries %>% 
    filter(job_title == input$job_title) %>%
    summarise(avg_salary = mean(salary_in_usd, na.rm = TRUE)) %>%
    pull(avg_salary)
  
  gauge(avg_salary, min = 0, max = max(salaries$salary_in_usd), 
        gaugeSectors(success = c(0, avg_salary/2), warning = c(avg_salary/2, avg_salary), danger = c(avg_salary, max(salaries$salary_in_usd))))
})
```

Column {data-width=350}
-------------------------------------
Median Salary

```{r}
renderGauge({
  median_salary <- salaries %>% 
    filter(job_title == input$job_title) %>%
    summarise(median_salary = median(salary_in_usd, na.rm = TRUE)) %>%
    pull(median_salary)
  
  gauge(median_salary, min = 0, max = max(salaries$salary_in_usd), 
        gaugeSectors(success = c(0, median_salary/2), warning = c(median_salary/2, median_salary), danger = c(median_salary, max(salaries$salary_in_usd))))
})

```

Map
=======================================================================
Column {data-width=350}
-------------------------------------

```{r}
selectInput("job_title", "Job Title:", 
            choices = unique(salaries$job_title), 
            selected = unique(salaries$job_title)[1])

```


```{r}
renderLeaflet({
  map_data <- salaries %>%
    filter(job_title == input$job_title_map) %>%
    group_by(company_location) %>%
    summarize(avg_salary = mean(salary_in_usd, na.rm = TRUE)) %>%
    left_join(country_coords, by = c("company_location" = "country"))
  
  leaflet(map_data) %>%
    addTiles() %>%
    addCircles(lng = ~lon, lat = ~lat, weight = 1,
               radius = ~sqrt(avg_salary) * 500,
               popup = ~paste(company_location, "<br>", "Average Salary: $", round(avg_salary, 2)))
})

```


About
=======================================================================

This website anonymously gathers salary data from professionals worldwide in the fields of AI, ML, and Data Science, making the information publicly accessible for anyone to use, share, and explore.

The main objective is to provide data that offers better insights into global salary trends. This helps new entrants, seasoned professionals, hiring managers, recruiters, startup founders, and individuals considering a career change make well-informed decisions.

## Dataset Description

The dataset includes 11 columns with the following details:

- **work_year**: The year in which the salary was paid.
- **experience_level**: The job experience level during the year, categorized as:
    - **EN**: Entry-level / Junior
    - **MI**: Mid-level / Intermediate
    - **SE**: Senior-level / Expert
    - **EX**: Executive-level / Director
- **employment_type**: The employment type for the role, including:
    - **PT**: Part-time
    - **FT**: Full-time
    - **CT**: Contract
    - **FL**: Freelance
- **job_title**: The job title during the year.
- **salary**: The total gross salary amount paid.
- **salary_currency**: The currency of the salary paid, represented as an ISO 4217 currency code.
- **salary_in_usd**: The salary converted to USD, based on the average exchange rate for the respective year via fxdata.foorilla.com.
- **employee_residence**: The primary country of residence of the employee during the work year, represented as an ISO 3166 country code.
- **remote_ratio**: The proportion of work done remotely, categorized as:
    - **0**: No remote work (less than 20%)
    - **50**: Partially remote/hybrid
    - **100**: Fully remote (more than 80%)
- **company_location**: The country of the employer's main office or contracting branch, represented as an ISO 3166 country code.
- **company_size**: The average number of employees at the company during the year, categorized as:
    - **S**: Small (less than 50 employees)
    - **M**: Medium (50 to 250 employees)
    - **L**: Large (more than 250 employees)

